State of the Practice for Lattice Boltzmann Method Software

### 2022-04-25 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Revisions to Section 3

We use ten research questions to structure and focus our discussion. Our
methodology \citep{SmithEtAl2021, SmithAndMichalski2022} is driven by these
questions. For each question below, we point to the section that contains our
answer.  We start with identifying the examples of LBM software where we have
access to the source code:

@@

All 24 packages are installable on a Unix-like system. Eight packages are also
installable on Windows, and five on macOS. Twenty packages list their operating
system requirements, but only four (ESPResSo, LUMA, Mechsys, Sailfish) list
compatible operating system versions. Ubuntu provided the testing platform for
all but one of the packages. TCLB was tested on CentOS, as suggested by its
installation instructions.

All but one of the software packages (LatBo.jl) have automated at least part of
the installation process. Most of these packages, such as waLBerla and
SunlightLB, use Make to automate the installation, and a few of them, like
lbmpy, use custom scripts.

We quickly fixed errors encountered during the installation process, thanks to
descriptive error messages. Systems that provided vague error messages, such as
messages that did not specify which action or file was at fault, were harder to
troubleshoot. Only three software packages (HemeLB, LB3D, lbmpy) that displayed
a descriptive error message were not recoverable, and most of these instances
were due to hardware and operating system incompatibility, such as the
requirement of CUDA. Fourteen software packages broke during installation. Some
packages, such as LB2D-Prime and LB3D-Prime did not provide a definitive message
of the success or failure of the installation. In these instances, validating
the installation required performing a tutorial or running a script, as
described below, if these were available.

About half of the installation instructions assume the installation environment
has not already installed dependent packages. Unfortunately, software packages,
like \esp, Ludwig and LUMA, frequently don't list all their dependencies.
Sometimes only an error message during the installation process informs the user
of the requirement for additional packages. In case like this, we suggest a
detailed rewrite of the installation instructions to accommodate a clean
operating system. A virtual machine provides such an environment for testing
purposes.

Seventeen software packages require installation of less than ten dependencies.
All but one software package (LatBo.jl) require less than 20 dependencies. Some
packages may automatically install additional dependencies in the background.
Twenty of the software packages do not explicitly indicate the versions of
software dependencies. Some software package installation issues could have been
avoided by providing a list of dependencies. Sixteen software packages do not
have detailed instructions for installing dependencies. Sixteen software
packages have less than ten manual installation steps. If one command installs
dependencies, then none of the software packages take more than 20 steps to
install. The average number of steps is eight, and the fewest is two
(LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages. Software packages with a higher score (ESPResSo, MechSys, OpenLB,
DL\_MESO, Palabos) tend to have one set of linear installation instructions
written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for installing dependencies. The top-ranked packages often
incorporate some automation of the installation process. The number of
dependencies a package has does not correlate with a higher score. The ability
to validate the installation process, say through tutorials or test cases, leads
to a higher score. Apparently active work on a project improves installability
since the top seven ranked packages are all classified as alive.

@@

Tutorials are available for 19 of the software packages. Generally they are
linearly written and easy to follow. However, only eight tutorials provide an
expected output. It is not possible to verify the correctness of the output of
the software packages that are missing this key information. In these cases the
user may need to assume correctness if there are no visible errors.
\href{https://www.walberla.net/doxygen/index.html} {waLBerla} provides a
particularly detailed tutorial, including expected output.

@@

The analysis of surface reliability focused on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, and syntax errors in code files. In some instances the error
messages were vague. Several automatic installation processes could not find or
load dependencies. In these instances the installation tried to access outdated
external repositories. We recovered and verified seven of the installations,
and we assumed that one of the installations (LB3D-Prime) was recovered due to the
absence of any way to verify otherwise. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
ranked packages are alive.

@@

Testing includes handling unexpected input, such as incorrect data types, empty
input, and missing files or links. A reasonable response, including appropriate
error messages and an absence of unrecoverable system failures, means success.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They output descriptive error messages or at least
did not crash. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
runtime robustness. All successfully installed software packages that require a
plain text input file correctly handled an unexpected change to these input
files, including a replacement of new lines with carriage returns. Furthermore,
nine of the top 10 ranked packages are alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, the packages focus on varied CFD problems, with varying parameters,
and are technically different from each other (as shown in
Table~\ref{tbl_features}). Due to this, a comparison of run-time performance is
not appropriate. Instead, we looked through each software package's artifacts
for evidence of considering performance. The artifacts of 18 software packages
mentioned parallelization. This included GPU processing and the CUDA parallel
computing platform, which were mentioned in the artifacts of six packages
(ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). When a high degree of
parallelization is possible, GPUs provide superior speed compared to CPUs. The
software package TCLB is implemented in a highly efficient multi-GPU code to
achieve performance suitable for model optimization \citep{rutkowski2020open}.
In the Ludwig package uses a so-called mixed mode approach implementing
fine-grained parallelism on the GPU.  MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. In general, the packages
that require GPU and CUDA have better performance at the expense of
installability and surface reliability.

@@

We reviewed software package artifacts for the presence of a tutorial, a user
manual, documented user characteristics, and a user support model. In total 19
software packages have a tutorial, 13 have a user manual, and 11 have both. The
tutorials vary in scope and substance, and eight include expected output. Most
user manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Five
software packages (laboetie, LIMBES, Ludwig, Musubi, Palabos) document expected
user characteristics. Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All but one
of the packages (LIMBES) have a user support model, and many of them have
multiple avenues of user support. The most popular avenue of support is Git,
followed by email and forums. One software package (OpenLB) has an FAQ page.

@@

We reviewed software packages for the presence of artifacts and recorded every
type of artifact that is not a code file. (The list of artifacts will be
discussed further in Section~\ref{Sec_CompareArtifacts}.)  We also reviewed the
software packages for software release and documentation version numbers. This
information can be used to troubleshoot issues and organize documentation. All
but three software packages (LatBo.jl, LB3D-Prime, MechSys) have source code
release and documentation version numbers.

When present, we noted how the team reviews code, and how users can contribute
to the project. In total, 11 software packages have this information, which was
found in various artifacts, including in developer guides, contributor guides,
user guides, developer webpages, and README files.

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50 percent of their issues closed. All the top five overall ranked packages
have most of their issues closed. Alive packages (11 use issue tracking) have
64\% of their issues closed, while dead packages (3 use issue tracking) have
71\% of their issues closed. Table~\ref{gitrepodata} presents this information.
With respect to the host for the git repository, 13 packages use GitHub and two
(Palabos, waLBerla) use GitLab. Of the other packages, one package (SunlightLB)
uses CVS for issue tracking and version control, and seven packages do not
appear to use any issue tracking system.

@@

Table~\ref{gitrepodata} shows the percentage of code that is comments, in
decreasing order. Maintainability improves for packages with a
higher percentage of comments. Comments represent more than 10 percent of code
files in 16 packages, and the average percentage of code comments is about 14
percent. All the top five overall ranked packages have more than the average.
LUMA has only 0.2 percent comments, the fewest of any package. This package has
the most lines of source code, with over four million. The next largest package
is \esp{} with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Maintainable packages by our ranking have more than 10 percent of the
code as comments. Four of the top five ranked packages are alive.

@@

We measured the total number of source code files for each project. We assume
that numerous source files increases reusability, since more files implies
increased modularization. Some packages have more features than others.  This is
assumed to contribute to reusability, since they have more source code for
potential reuse. We reviewed the software packages for the presence of API
documentation, which suggests that the developers considered potential future
reuse.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. However, these packages do have an abundance of reusable
components. Nine of the top 10 ranked packages are alive.

@@

The package waLBerla scored high on reusability because of its focus on
modularity.  The modularity in the waLBerla framework enhances productivity,
reusability, and maintainability \citep{bauer2021walberla}. The design of
waLBerla has enabled it to be successfully applied in several projects as a
basis for various extensions \citep{bauer2021walberla}.

@@

To measure understandability, we reviewed ten random source code files per
package.  By using only 10 files we may incorrectly estimate understandability,
but for practical consideration our time measuring each package has to be
limited.

All the packages appear to have consistent indentation and formatting. Only
HemeLB, LUMA, and Musubi explicitly identify coding standards that are used
during development. Generally, the software packages use consistent,
distinctive, and meaningful code identifiers. Only four packages (LB2D-Prime,
LB3D-Prime, LIMBES, MP-LABS) appear to use vague identifiers, such as single
letters for variables. Thirteen source code files used symbolic constants for
various parameters, mathematical constants, and matrix definitions. The code
includes appropriate comments for all packages, with the comments clearly
indicating what is being done (as opposed to how it is being done). The source
code of 11 packages includes notes on domain algorithms. Table~\ref{moduledata}
suggests that the software packages are modular to various degrees. When
observing the source code files, 14 of the packages have a consistent style and
order of function parameters.

@@

We reviewed the software package artifacts for the identification of a specific
development model, like a waterfall or agile development model, and the presence
of documentation recording the development process and standards. They were also
reviewed for the identification of the development environment, and the presence
of release notes. The packages tended to not explicitly use well-known
development models. This was also noted in the interviews with developers
(Section~\ref{Sec_CompareMethodologies}). The development teams of these
packages are fairly small and easily organized without the need for such
processes. Seven of the software packages did have some artifacts outlining the
general development process, how to contribute, and the status of the package or
its components. Eight of the packages have artifacts that note the development
environment. While this information could help developers, and would improve
transparency, the small close-knit nature of the development teams make
explicitly specifying this information practically unnecessary. Nine of the
software packages included version release notes.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access notes accompanying software releases. Four of the
top five ranked packages are alive.

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking of the software
packages. In the absence of other information on priorities, we calculated the
overall ranking by assuming an equal weighting between all qualities. If we
weigh the qualities differently, the overall software package ranking will
change. Software packages with an overall high score ranked high in at least
several of the individual qualities that were quantitatively measured.

@@

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{1.25cm}p{1.75cm}p{1.75cm}p{1.75cm}p{1.75cm} }
			\toprule
			Name & Our Ranking & Repository Stars & Repository Star Rank &
			Repository Watches & Repository Watch Rank\\
			\midrule
			ESPResSo & 1 & 145 & 2 & 19& 2\\
			Ludwig & 2 & 27 & 8 & 6 & 7\\
			Palabos & 3 & 34 & 6 & GitLab & GitLab\\
			OpenLB & 4 & N/A & N/A & N/A& N/A\\
			LUMA & 5 & 33 & 7 & 12& 4\\
			pyLBM & 6 & 95 & 3 & 10& 5\\
			DL\_MESO (LBE) & 7 & N/A & N/A & N/A & N/A\\
			Musubi & 8 & N/A & N/A & N/A & N/A\\
			Sailfish & 9 & 186 & 1 & 41& 1\\
			waLBerla & 10 & 20 & 9 & GitLab & GitLab\\
			laboetie & 11 & 4 & 13 & 5 & 8\\
			TCLB & 12 & 95 & 3 & 16 & 3\\
			MechSys & 13 & N/A & N/A & N/A & N/A\\
			lettuce & 14 & 48 & 4 & 5 & 8\\
			\esp & 15 & 35 & 5 & 12 & 4\\
			MP-LABS & 16 & 12 & 11 & 2 & 9\\			
			SunlightLB & 17 & N/A & N/A & N/A & N/A\\
			LB3D & 18 & N/A & N/A & N/A & N/A\\			
			LIMBES & 19 & N/A & N/A & N/A & N/A\\
			LB2D-Prime & 20 & N/A & N/A & N/A & N/A\\		
			HemeLB & 21 & 12 & 11 & 12 & 4\\
			lbmpy & 22 &  11 & 12 & 2 & 9\\	
			LB3D-Prime & 23 & N/A & N/A & N/A & N/A\\	
			LatBo.jl & 24 & 17 & 10 & 8 & 6\\			
			\bottomrule
		\end{tabular}
		\caption{Repository Ranking Metrics (Sorted by Our Ranking)} \label{repometrics}
	\end{center}
\end{table}

>>>>

We use 10research questions to structure and focus our discussion. Our
methodology \citep{SmithEtAl2021, SmithAndMichalski2022} is driven by these
questions. For each question below, we point to the section that contains our
answer.  We start with identifying the examples of LBM software where we have
access to the source code:

@@

All 24 packages are installable on a Unix-like system. Eight packages are also
installable on Windows, and five on macOS. Twenty packages list their operating
system requirements, but only four (ESPResSo, LUMA, Mechsys, Sailfish) list
compatible operating system versions. Ubuntu provided the testing platform for
all packages, except for TCLB, which we tested on CentOS, as suggested by its
installation instructions.

All but one of the software packages (LatBo.jl) have automated at least part of
the installation process. Most packages, such as waLBerla and SunlightLB, use
Make to automate the installation, and a few of them, like lbmpy, use custom
scripts.

In most of the cases with installation errors, descriptive error messages
allowed us to proceed. Systems that provided vague error messages, such as
messages that did not specify which action or file was at fault, were harder to
troubleshoot. Only three software packages (HemeLB, LB3D, lbmpy) that displayed
a descriptive error message were not recoverable, and most of these instances
were due to hardware and operating system incompatibility, such as the
requirement of CUDA. Fourteen software packages broke during installation. Some
packages, such as LB2D-Prime and LB3D-Prime did not provide a definitive message
of the success or failure of the installation. In these instances, validating
the installation required performing a tutorial or running a script, if these
were available.

About half of the installation instructions assume the user is already aware of,
and has already installed, dependent packages. Many packages, like \esp, Ludwig
and LUMA, do not list all their dependencies. Sometimes only an error message
during the installation process informs the user of a required dependency. In
cases like this, we suggest a detailed rewrite of the installation instructions
to accommodate a clean operating system. A virtual machine can provide an
environment for installability testing.

Seventeen software packages require installation of less than 10 dependencies.
All software package (except for LatBo.jl) require less than 20 dependencies.
Some packages may automatically install additional dependencies in the
background. Twenty of the software packages do not explicitly indicate the
versions of software dependencies. Some software package installation issues
could have been avoided by providing a list of dependencies. Sixteen software
packages do not have detailed instructions for installing dependencies. Sixteen
software packages have less than 10 manual installation steps. If one command
installs dependencies, then none of the software packages take more than 20
steps to install. The average number of steps is eight, and the fewest is two
(LB3D-Prime).

@@

Figure~\ref{Fig_Installability} shows the installability ranking of the software
packages. Software packages with a higher score (ESPResSo, MechSys, OpenLB,
DL\_MESO, Palabos) tend to have one set of linear installation instructions,
written as if the person doing the installation has none of the dependencies
installed. The instructions often list compatible operating system versions and
include instructions for installing dependencies. The top-ranked packages often
incorporate some automation of the installation process. The number of
dependencies a package has does not correlate with a higher score. The ability
to validate the installation process, say through tutorials or test cases, leads
to a higher score. Apparently active work on a project improves installability
since the top seven ranked packages are all classified as alive.

@@

Tutorials are available for 19 of the packages. Generally they are linearly
written and easy to follow. However, only eight tutorials provide expected
outputs, making it difficult to verify the correctness of their output. In these
cases the user may need to assume correctness if there are no visible errors.
\href{https://www.walberla.net/doxygen/index.html} {waLBerla} provides a
particularly detailed tutorial, including expected output.

@@

The analysis of surface reliability focuses on package installation and
tutorials. Errors occurred when installing 16 of the software packages. Every
instance prompted an error message. These messages indicated unrecognized
commands (even when following the installation guide), missing links, missing
dependencies, and syntax errors in code files. In some instances the error
messages were vague. Several automatic installation processes could not find or
load dependencies. In these instances the installation tried to access outdated
external repositories. We recovered and verified seven of the installations, and
we assumed that one of the installations (LB3D-Prime) was recovered, since there
was no way to test the installation. The installation of eight of the software
packages could not be recovered. Most of these broken installations could not
find external dependencies, encountered system incompatibilities, or displayed
vague error messages.

@@

Figure~\ref{Fig_Reliability} shows the surface reliability ranking of the
software packages. Software packages with a high score either did not break
during installation, or the broken installation was recoverable. All the top
five ranked packages have tutorials. The package pyLBM broke during tutorial
testing, but a descriptive error message helped in recovery. Nine of the top 10
packages are alive.

@@

Robustness testing includes handling unexpected input, such as incorrect data
types, empty input, and missing files or links. A reasonable response, including
appropriate error messages and an absence of unrecoverable system failures,
means success.

@@

Figure~\ref{Fig_Robustness} shows the surface robustness ranking of the software
packages. Software packages with a high score behaved reasonably in response to
unexpected input as described above. All the software packages that installed
correctly passed this test. They did not crash and usually output descriptive
error messages. Software packages with a lower surface robustness score did not
install correctly, so their robustness score may not be a true reflection of
their runtime robustness. All successfully installed software packages that
require a plain text input file correctly handled an unexpected change to these
input files, including a replacement of new lines with carriage returns.
Following the trend of the other qualities, nine of the top 10 ranked packages
are alive.

@@

Although the software packages all apply LBM to solve scientific computing
problems, they differ in scope and capabilities, as shown in
Table~\ref{tbl_features}. Therefore, tests to compare run-time performance are
not appropriate. Instead, we look through each software package's artifacts
for evidence of considering performance. The artifacts of 18 software packages
mentioned parallelization. This included GPU processing and the CUDA parallel
computing platform, which were mentioned in the artifacts of six packages
(ESPResSo, lbmpy, lettuce, pyLBM, Sailfish, TCLB). When a high degree of
parallelization is possible, GPUs provide superior speed compared to CPUs. The
software package TCLB is implemented in a highly efficient multi-GPU code to
achieve performance suitable for model optimization \citep{rutkowski2020open}.
The Ludwig package uses a so-called mixed mode approach implementing
fine-grained parallelism on the GPU.  MPI is used for even larger scale
parallelism \citep{gray2013ludwig}. While one software package (Sailfish)
required CUDA and GPU processing, some (ESPResSo, lbmpy, lettuce, pyLBM, TCLB)
have the option of using either the GPU or the CPU. In general, the packages
that require GPU and CUDA have better performance at the expense of
installability and surface reliability.

@@

We review software repositories for the presence of a tutorial, a user manual,
documented user characteristics, and a user support model. In total 19 software
packages have a tutorial, 13 have a user manual, and 11 have both. The tutorials
vary in scope and substance, and eight include expected output. Most user
manuals are in the form of a file that can be downloaded, while some are
rendered on a webpage. Some packages (waLBerla) do not have a user manual, but
do have useful documentation distributed throughout their webpages. Five
software packages (laboetie, LIMBES, Ludwig, Musubi, Palabos) document expected
user characteristics. Users are typically scientists or engineers. Their
background is often physics, chemistry, biophysics, or mathematics. All projects
(except LIMBES) have a user support model, and many of them have multiple
avenues of user support. The most popular avenue of support is Git, followed by
email and forums. One software package (OpenLB) has an FAQ page.

@@

We review software packages for the presence of artifacts and recorded every
type of artifact that is not a code file. (The artifacts will be discussed
further in Section~\ref{Sec_CompareArtifacts}.)  We also reviewed the software
packages for software release and documentation version numbers. This
information can be used to troubleshoot issues and organize documentation. All
but three software packages (LatBo.jl, LB3D-Prime, MechSys) have source code
release and documentation version numbers.

We look for evidence that teams review code and have information on how users
can contribute. In total, 11 software packages have this information, found in
various artifacts, including in developer guides, contributor guides, user
guides, developer webpages, and README files.

Twenty-three packages use issue tracking, 15 of which use GitHub or GitLab to
host their git repository, seven use email, and one (SunlightLB) uses
SourceForge. Most software packages that use GitHub or GitLab have the majority
of their issues closed, and only three (laboetie, lettuce, Sailfish) have less
than 50\% of their issues closed. All the top five ranked packages have most of
their issues closed. Alive packages (11 use issue tracking) have 64\% of their
issues closed, while dead packages (3 use issue tracking) have 71\% of their
issues closed. Table~\ref{gitrepodata} presents this information. With respect
to the host for the git repository, 13 packages use GitHub and two (Palabos,
waLBerla) use GitLab. Of the other packages, one package (SunlightLB) uses CVS
for issue tracking and version control, and seven do not appear to use any issue
tracking system.

@@

Table~\ref{gitrepodata} shows the percentage of code that is comments, in
decreasing order. We assume maintainability improves for packages with a higher
percentage of comments. Comments represent more than 10\% of code files in 16
packages, and the average percentage of code comments is about 14\%. All the top
five overall ranked packages have more than the average. LUMA has only 0.2\%
comments, the fewest of any package. However, this package also has the most
lines of source code, with over four million. The next largest package is \esp{}
with one million.

@@

Figure~\ref{Fig_Maintainability} shows the maintainability ranking of the
software packages. Software packages with a high score provide version numbers
on documents and source code releases, have an abundance of high quality
artifacts, and use an issue tracking tool and version control system. These
packages also appear to manage their issue tracking, having most of their issues
closed. Maintainable packages by our ranking have more than 10\% of the code as
comments. Four of the top five ranked packages are alive.

@@

We measure the total number of source code files for each project. We assume
that numerous source files increases reusability, since more files implies
increased modularization. Some packages have more features than others.  We
assume this contributes to reusability, since they have more source code for
potential reuse. We reviewed the software packages for the presence of API
documentation, which suggests that the developers considered potential future
reuse.

@@

Figure~\ref{Fig_Reusabilty} shows the reusability ranking of the software
packages. Software packages with a high score have thousands of source code
files and API documentation. The highest scoring packages, ESPResSo and
waLBerla, have extensive functionality, including graphical visualizations and
non-LBM modelling. For this reason, a comparison with other software packages is
not on a level field. Nine of the top 10 ranked packages are alive.

@@

The package waLBerla scored high on reusability because of its focus on
modularity.  The modularity in the waLBerla framework enhances productivity,
reusability, and maintainability \citep{bauer2021walberla}. The design of
waLBerla has enabled it to be successfully applied as a basis for extension in
several projects \citep{bauer2021walberla}.

@@

To measure understandability, we review 10 random source code files per package.
Using only 10 files may incorrectly estimate understandability, but for
practical considerations we have to limit our time measuring each package.

All the packages appear to have consistent indentation and formatting. Only
HemeLB, LUMA, and Musubi explicitly identify coding standards that are used
during development. Generally, the software packages use consistent,
distinctive, and meaningful code identifiers. Only four packages (LB2D-Prime,
LB3D-Prime, LIMBES, MP-LABS) appear to use vague identifiers, such as single
letters for variables. Thirteen source code files use symbolic constants for
various parameters, mathematical constants, and matrix definitions. The code
includes appropriate comments for all packages, with the comments clearly
indicating what is being done (as opposed to how it is being done). The source
code of 11 packages include notes on domain algorithms. Table~\ref{moduledata}
suggests that the software packages are modular to various degrees. When
observing the source code files, 14 of the packages have a consistent style and
order of function parameters.

@@

We review the software artifacts for process documentation, a development
environment description, and release notes.  The packages tend to not explicitly
use well-known development models (see also
Section~\ref{Sec_CompareMethodologies}). The development teams of these packages
are fairly small and easily organized without the need for such processes. Seven
of the software packages did have some artifacts outlining the general
development process, how to contribute, and the status of the package or its
components. Eight of the packages have artifacts that note the development
environment. While this information could help developers, and would improve
transparency, the small close-knit nature of the development teams make
explicitly specifying this information practically unnecessary. Nine of the
software packages included version release notes.

@@

Figure~\ref{Fig_VisibilityTransparency} shows the visibility and transparency
ranking of the software packages. Software packages with a high score have an
explicit development model and defined development process. They also have
detailed and easy to access software release notes. Four of the top five ranked
packages are alive.

@@

Figure~\ref{Fig_OverallScore} shows the overall ranking. In the absence of other
information on priorities, we calculated the ranking by assuming an equal
weighting between all qualities. Software packages with an overall high score
ranked high in at least several of the individual qualities.

@@

\begin{table}[ht!]
	\begin{center}
		\begin{tabular}{ p{3cm}p{1.25cm}p{1.75cm}p{1.75cm}p{1.75cm}p{1.75cm} }
			\toprule
			Name & Our Ranking & Repository Stars & Repository Star Rank &
			Repository Watches & Repository Watch Rank\\
			\midrule
			ESPResSo & 1 & 145 & 2 & 19& 2\\
			Ludwig & 2 & 27 & 8 & 6 & 7\\
			Palabos & 3 & 34 & 6 & GitLab & GitLab\\
			OpenLB & 4 & N/A & N/A & N/A& N/A\\
			LUMA & 5 & 33 & 7 & 12 & 4\\
			pyLBM & 6 & 95 & 3 & 10 & 5\\
			DL\_MESO (LBE) & 7 & N/A & N/A & N/A & N/A\\
			Musubi & 8 & N/A & N/A & N/A & N/A\\
			Sailfish & 9 & 186 & 1 & 41& 1\\
			waLBerla & 10 & 20 & 9 & GitLab & GitLab\\
			laboetie & 11 & 4 & 13 & 5 & 8\\
			TCLB & 12 & 95 & 3 & 16 & 3\\
			MechSys & 13 & N/A & N/A & N/A & N/A\\
			lettuce & 14 & 48 & 4 & 5 & 8\\
			\esp & 15 & 35 & 5 & 12 & 4\\
			MP-LABS & 16 & 12 & 11 & 2 & 9\\			
			SunlightLB & 17 & N/A & N/A & N/A & N/A\\
			LB3D & 18 & N/A & N/A & N/A & N/A\\			
			LIMBES & 19 & N/A & N/A & N/A & N/A\\
			LB2D-Prime & 20 & N/A & N/A & N/A & N/A\\		
			HemeLB & 21 & 12 & 11 & 12 & 4\\
			lbmpy & 22 &  11 & 12 & 2 & 9\\	
			LB3D-Prime & 23 & N/A & N/A & N/A & N/A\\	
			LatBo.jl & 24 & 17 & 10 & 8 & 6\\			
			\bottomrule
		\end{tabular}
		\caption{Repository Ranking Metrics (Sorted by Our Ranking)} \label{repometrics}
	\end{center}
\end{table}
